---
output:
  github_document
---
# mfbvar
[![Build Status](https://travis-ci.org/ankargren/mfbvar.svg?branch=master)](https://travis-ci.org/ankargren/mfbvar)
[![](http://www.r-pkg.org/badges/version/mfbvar)](http://www.r-pkg.org/pkg/mfbvar)
[![Coverage status](https://codecov.io/gh/ankargren/mfbvar/branch/master/graph/badge.svg)](https://codecov.io/github/ankargren/mfbvar?branch=master)

## Overview 
The `mfbvar` package implements a steady-state prior and a Minnesota prior for state space-based mixed-frequency VAR models.
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
thm = knitr::knit_theme$get("edit-xcode")
knitr::knit_theme$set(thm)
```

### Prior specification
First, obtain some data stored in the package.
```{r}
library(mfbvar)
Y <- mf_sweden
head(Y)
tail(Y)
```
Next, we create a minimal prior object. We must specify: 1) data, 2) the frequency of the data, 3) the number of lags, 4) the length of burn-in and main chains, respectively. This is done by calling the `set_prior()` function and giving named arguments. The resulting object is of class `mfbvar_prior` and has a basic `print` method.
```{r}
prior_obj <- set_prior(Y = Y, freq = c(rep("m", 4), "q"), 
                       n_lags = 4, n_burnin = 1000, n_reps = 1000)
```
Warnings are produced because we haven't specified values for some of the prior elements and instead the function uses default values.

There is a print method for the prior object, showing some basic information:
```{r}
prior_obj
```
The message tells us what elements of the prior have not yet been set, and if each of the two priors can be run with the current specification. The check is very minimal; the steady-state prior cannot be used to make forecasts (which it will attempt to if `n_fcst` is greater than `0`) unless also `d_fcst` is given, but to run the model with no forecasts only the three indicated elements are missing.

The summary method provides a little bit more detail:
```{r}
summary(prior_obj)
```

### Model estimation
As the print method told us before, we can run the Minnesota prior, but not the steady-state prior with the current prior specification. The model is estimated by calling `estimate_mfbvar()`.
```{r, cache = TRUE}
mod_minn <- estimate_mfbvar(mfbvar_prior = prior_obj, prior_type = "minn")
```

To use the steady-state prior, we need to specify `d`, `prior_psi_mean` and `prior_psi_Omega`. We specify the prior moments for $\psi$ using the helper function `interval_to_moments()` which converts 95 % prior probability intervals to prior moments, assuming independence.
```{r}
prior_intervals <- matrix(c( 6,   7,
                             0.1, 0.2,
                             0,   0.5,
                            -0.5, 0.5,
                             0.4, 0.6), ncol = 2, byrow = TRUE)
psi_moments <- interval_to_moments(prior_intervals)
prior_psi_mean <- psi_moments$prior_psi_mean
prior_psi_Omega <- psi_moments$prior_psi_Omega
```

Instead of creating a new prior object, we can update the old by use of the `update_prior()` function. Note also that it is possible to specify `"intercept"` for `d` rather than a matrix containing a constant for the deterministic term.
```{r}
prior_obj <- update_prior(prior_obj, d = "intercept", prior_psi_mean = prior_psi_mean, 
                          prior_psi_Omega = prior_psi_Omega)
prior_obj
```

It is now possible to estimate the model using the steady-state prior.
```{r, cache = TRUE}
mod_ss <- estimate_mfbvar(prior_obj, "ss")
```

It is also allowed to temporarily override elements in the prior object by adding them as separate arguments to the `estimate_mfbvar()` function. Thus, to get forecasts eight steps ahead we would use:
```{r, cache = TRUE}
mod_minn <- estimate_mfbvar(prior_obj, "minn", n_fcst = 8)
mod_ss <- estimate_mfbvar(prior_obj, "ss", n_fcst = 8)
```

### Processing results
The resulting objects contain all of the posterior information. The returned objects from `estimate_mfbvar()` are of class `mfbvar` and `mfbvar_ss` or `mfbvar_minn`.
```{r}
class(mod_minn)
class(mod_ss)
```

For forecasts, there is a `predict` method for class `mfbvar` which computes forecasts for selected quantiles. By default, it returns the 10%, 50% and 90% quantiles.
```{r}
predict(mod_minn, pred_quantiles = 0.5)
```
If desired, it can be requested in a tidy format.
```{r}
head(predict(mod_minn, pred_quantiles = 0.5, tidy = TRUE))
```

Calling plot on `mfbvar_ss` or `mfbvar_minn` objects produces plots of the forecasts and, by default, `5*n_fcst` of the preceding values.

```{r plot_minn, cache = TRUE}
plot(mod_minn)
``` 

The axis tick labels are too long and overlap. The `plot()` method returns a `ggplot`. Hence, modifying the plot simply amounts to adding layers in the usual `ggplot2` way. The method also allows for changing where the plot should begin.
```{r plot_ss, cache = TRUE}
library(ggplot2)
plot(mod_ss, plot_start = 1) +
  theme(axis.text.x = element_text(angle = 90))
```

There are also some basic `print` and `summary` methods for the two classes implemented.
```{r}
mod_minn
mod_ss
summary(mod_minn)
summary(mod_ss)
```
### Marginal data density estimation
To estimate the marginal data density, there is a generic function `mdd()` for which there are methods for classes `mfbvar_ss` and `mfbvar_minn`.
```{r, cache = TRUE}
mdd_minn <- mdd(mod_minn)
mdd_ss_1 <- mdd(mod_ss)
mdd_ss_2 <- mdd(mod_ss, p_trunc = 0.5)

mdd_minn
mdd_ss_1
mdd_ss_2
```
